Handling String-Number Conversion in String Constraint Solving

### 2019-11-14 draft.tex

\newcommand{\nat}{\mathbb{N}}
\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\yfc}[1]{{\color{blue}YFC: #1}}
\newcommand{\petr}[1]{{\color{pink}Petr: #1}}
\newcommand{\sti}[1]{\mbox{\textsf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
%\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvarone}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvartwo}{\mbox{$V'_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{V}
\newcommand{\cvartwo}{V'}
\newcommand{\cvar}{V}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}[1]{\mathit{decode_{#1}}}
\newcommand{\parikhfof}[1]{\Phi_{\parikhof{#1}}}
\newcommand{\pim}{I_{\#}}
%\newcommand{\syncop}{\Cap}
\newcommand{\syncop}{\times}
\newcommand{\syncof}[2]{#1 \syncop #2}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncT}{T_\syncop}
\newcommand{\pvarsof}[1]{\#{#1}}
\newcommand{\pvar}{\pvarsof V}
\newcommand{\pvarone}{\pvarsof V}
\newcommand{\pvartwo}{\pvarsof {V'}}
\newcommand{\defeq}{::=}

@@

Through the paper, we will use quantifier-free linear integer arithmetic formulas, and call them \emph{linear formulas} for short.   
Given a linear formula $\phi$, a set of variables $V$ that includes all variables occurring in $\phi$, and an \emph{interpretation} of $V$, i.e., a function $I:V\rightarrow \nat$, 
we denote by $I\models_V \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner), and call $I$ a \emph{model} of $\phi$ over $V$.
We use $\modelsof \phi$ to denote the set all models of $\phi$ over $V$. 
We drop the reference to $V$ and write just ``model of $\phi$'',  $\modelsof \phi$, and $\models$ when no confusion may arise. For a set of interpretations $S$, we use $S_\Sigma$ to denote the set of interpretations obtained from $S$ by projecting the domain to $\Sigma$.

@@

The crucial feature of parametric flat automata is that their semantics can be faithfully represented by a linear formula and handled efficiently by an SMT solver. 
\lh{, allowing to circumvent more costly automata algorithms and decision procedures. Concretize and justify this better?}\yfc{The two sentences below read a bit strange, maybe need to rephrase it.}
%
Achieving this was the purpose of introducing the structural constraints of PFA. 
Their implication central for our method is that every parametric word $x\in L(A)$ is uniquely determined by its Parikh image $\parikhof{x}$, as formalized by the following lemma.

\begin{lemma}\label{lemma:parikh}
For an PFA $A$ and $x,y\in L(A)$, if $\parikhof x = \parikhof y$ then $x = y$. 
\end{lemma}
%
Indeed, since every variable appears on at most one transition, then the $\parikhof x$ value of all variables appearing within the same cycle is the same, and it is equal to the number of repetitions of that cycle in the accepting run. 
%
The accepting run on $x$ (and so $x$ itself) can thus be reconstructed from $\parikhof x$. \lh{say precisely how?}
%
For example, in the automaton of Figure~\ref{fig:sfa_def}, 
from $\parikhwof x {v_0} = \parikhwof x {v_1} = \cdots = \parikhwof x {v_{m-1}}=1$ and $\parikhwof x {v^0_1}=\parikhwof x {v^1_1}=\parikhwof x {v^2_1}=2$ we derive that $x=v_0(v^0_1v^1_1v^2_1)^2v_1\cdots v_{m-1}$. 
We may say that $\parikhof x$ is an encoding of $x$, and define the decoding function $\decode A$ such that $\decode A(\parikhof x) = x$.

\begin{lemma}\label{lemma:decx}
For a PFA $A$, $L(A) = \{\decode A(\pim) \mid \pim\in\semof{\parikhfof A}\}$.
\end{lemma}

Since every word $w \in \semof{A}$ is obtained from a parametric word $x\in L(A)$ and an interpretation of character variables $I:\cvar\rightarrow\Sigma_\epsilon$ such that $w = I(x)$, and since $x = \decode A(\parikhof x)$, 
we may say that the assignment $I \cup \parikhof x$ uniting the interpretation $I$ and the Parikh image of the parametric word $x$ encodes $w$. Lemma~\ref{lemma:parikh} also implies that two different words are encoded by different united assignments (the Parikh image or the interpretation must be different), and hence we may extend the decoding function $\decode A$ to united assignments so that $\decode A(I \cup \parikhof x) = w$.
Encodings of words in $\semof{A}$ are encodings of parametric words of $L(A)$ paired with arbitrary interpretations of character variables $\cvar$, which are precisely all models of the formula $\parikhfof A$ interpreted over $\cvar\cup\pvarsof \cvar$, as formulated by the following corollary of Lemma~\ref{lemma:decx}:
%
\begin{corollary}\label{corollary:pfa}
For an PFA $A$ over character variables $V$,
$\semof{A} = \{\decode A(I\cup \pim) \mid I\cup\pim\in\semof{\parikhfof A}_{\cvar\cup\pvar}\}$.
\end{corollary}
%
A minor technical difficulty is that we wish to work with pure arithmetic formulas, 
but we may also wish to specify when a character variable is assigned $\epsilon$, which is not a number. 
Hence, to allow purely arithmetic formulas to talk about the value $\epsilon$, we encode $\epsilon$ as $-1$.
In the further development, $\epsilon$ will mean $-1$ when working with arithmetic formulas constraining interpretations of character variables.

\paragraph{Synchronization of PAs}
For two PAs $A$ and $A'$ over character variables $V$ and $V'$, respectively, 
it builds \emph{synchronization formula} $\syncfof A {A'}$, a linear formula over $\cvarone\cup \cvartwo \cup \pvarone\cup\pvartwo$ encoding the semantic intersection $\semof A \cap \semof {A'}$. The synchronization formula will be used in, e.g., the decision procedure of equality constraints $t_1=t_2$ to find a common interpretation of $t_1$ and $t_2$. More precisely, the models of the synchronization formula $\syncfof A {A'}$ correspond to the set of all pairs of word encodings $(I\cup \pim)\in \parikhof {A}$ and $(I'\cup \pim')\in\parikhof {A'}$ such that 
$\decode A(I\cup \pim) = \decode {A'}(I'\cup \pim')$, which constitute the intersection.

Particularly, a synchronization formula is build for two PAs  
$A = (Q,T,\cvarone,q_i,q_f)$ and ${A'} = (Q',T',\cvartwo,q_i',q_f')$ such that $\cvarone \cap \cvartwo = \emptyset$. 
%
It is extracted from the \emph{asynchronous product} of $A$ and $A'$. 
%
The asynchronous product is an automaton that uses $Q\times Q'$ as the set of states and 
$\cvarone_\epsilon\times\cvartwo_\epsilon$
as the alphabet. 
%
Every accepting run of $\syncof A {A'}$ corresponds to a pair of accepting runs, 
a run of $A$ over a parametric word $x$ and a run of ${A'}$ over a parametric word $x'$. 
The word accepted by the run of $\syncof A {A'}$ induces constraints on the interpretations of $I$ of $\cvarone$ and $I'$ of $\cvartwo$ under which the two parametric words have the same interpretation, i.e. $I(x) = I(x')$.

@@

The correctness of this construction is stated in Lemma~\ref{lemma:synccorrect} below. The lemma can also be understood as a precise specification the semantics of the synchronization formula, on which we rely for correctness of the construction of under-approximations of string constraints. 
%
\hide{
\begin{lemma}\label{lemma:synccorrect}
%$\semof {\exists V_\times . \syncfof A {A'}}\restrict{\cvarone \cup \cvartwo \cup \pvarsof{\cvarone} \cup \pvarsof\cvartwo} = \{(I\cup\pim)\cup(I'\cup\pim') \mid \decode A(I\cup\pim) = \decode {A'}(I'\cup\pim')\}$. 
$\semof {\syncfof A {A'}}_{\cvarone\cup \cvartwo \cup \pvarone\cup\pvartwo} = \{(I\cup\pim)\cup(I'\cup\pim') \mid \decode A(I\cup\pim) = \decode {A'}(I'\cup\pim')\}$. 
\end{lemma}
%
Another version:}
\begin{lemma}\label{lemma:synccorrect}
$
\semof {\syncfof A {A'}}_{\cvarone\cup \cvartwo \cup \pvarone\cup\pvartwo} = 
\{
(I\cup\parikhof{x})\cup(I'\cup\parikhof{x'}) \mid 
x\in L(A), x'\in L(A'), I(x) = I(x')
\}
$
\end{lemma}
Notice that if $A$ is flat (or, symmetrically, if $A'$ is flat), then $\semof A \cap \semof {A'}$ can be still decoded from $\semof {\syncfof A {A'}}$ due to Corollary~\ref{corollary:pfa}, 
namely in the form of $\semof A \cap \semof {A'} = \{\decode A(I\cup \pim) \mid (I\cup \pim) \in \semof{\syncfof A {A'}}_{\cvarone\cup \pvarone}\}$.

Informally, the lemma states that the models of $\syncfof A {A'}$ encode precisely the pairs of equivalent encodings of words from $\semof{A}$ and $\semof{A'}$, that constitute the intersection $\semof{A} \cap \semof{A'}$. A minor technical difficulty is that the modes of $\syncfof A {A'}$ include also an assignment to the auxiliary variables of $V_\times = (\cvarone_\epsilon\times\cvartwo_\epsilon) \cup \pvarsof{(\cvarone_\epsilon\times\cvartwo_\epsilon)}$. Hence in the lemma, we have to restrict the models to the original variables $\cvarone \cup \cvartwo \cup \pvarsof{\cvarone} \cup \pvarsof\cvartwo$. %This is  achieve by quantifying out the variables of $V_\times$.

%We actually have even something stronger, namely, that, $\syncfof A {A'}$ has all possible allignments. How do I say this ... or do I have to say this?
%\begin{lemma}
%For $\pim:\pvarone\rightarrow \nat$, $\pim':\pvartwo \rightarrow\nat$, $I:\cvarone\rightarrow\Sigma$, $I':\cvartwo\rightarrow\Sigma$, 
%we have $(I\cup\pim)\cup(I'\cup\pim')\in\semof {\syncfof A {A'}}$ if and only if $\decode A(I\cup\pim) = \decode {A'}(I'\cup\pim')$.
%\end{lemma}
In the rest of the section, we assume each string variable $x$ is restricted to a PFA $A_x$ with the set of character variables~$V_x$.

\yfc{we need to define \#{V} somewhere}

%\begin{lemma}
%	Symbolic flat automata are closed under concatenation.
%\end{lemma}

@@

\begin{itemize}
	\item $V' = \{ v_{(q,a,q')} \mid (q,a,q')\in T\}$
	\item $T' = \{ (q, v_{(q,a,q'), q') } \mid (q,a,q')\in T\}$
\end{itemize}
We use the linear formula $\Psi_c \defeq \bigwedge\limits_{(q,a,q')\in T} (v_{(q,a,q')} = a)$ to remember the relation between the fresh character variables in $V'$ and the symbol it represents. We add both $\syncfof {A_x }{A'} $ and $\Psi_c$ to $\Phi_O$. The relation between $A$ and $A'$ is expressed in Lemma~\ref{lemma:fa_pfa}.

\begin{lemma}\label{lemma:fa_pfa}
	$L(A)=	\{  I(x) \mid  I\in  \semof {\Psi_c} \wedge x \in L(A') \} $. 
\end{lemma}

The correctness of this procedure is expressed in Lemma~\ref{lemma:memcorrect} below. 
%
%\newcommand\restrict[1]{\raisebox{-.5ex}{$\big|$}_{#1}}
\begin{lemma}\label{lemma:memcorrect}
	$\semof {\syncfof {A_x }{A'} \wedge \Psi_c}_{V_x \cup \#V_x} = \{(I\cup\pim) \mid \decode {A_x}(I\cup\pim) \in L(A)\}$. 
\end{lemma}

Lemma~\ref{lemma:memcorrect} follows from Corollary~\ref{corollary:pfa}, Lemma~\ref{lemma:synccorrect}, and Lemma~\ref{lemma:fa_pfa}. Informally, the lemma says the models of the formula we add to $\Phi_O$, i.e. $\syncfof {A_x }{A'} \wedge \Psi_c$, characterize exactly the models of $x \in  L(A)$ under the restriction of $A_x$.

@@

Assume that the constraint we take from $\Phi_I$ is an equality constraint $\phi_e\defeq x_1\cdot x_2 \cdots x_n = x_{n+1}\cdot x_{n+2} \cdots x_m$. We first build two PFAs that encode the left-hand side term and the right-hand side term of the equality constraint. Intuitively, this can be achieved by connecting the PFAs of the string variables with epsilon transitions. However, there are two minor technical issues. First, our definition of PFA does not allow epsilon transition. Second, some string variable might occur multiple times in a equality constraint, but the definition of PFA allows each character variable occurs in only one transition.

@@

For the second problem, we rename the variables on the transitions of PFAs for different occurrences of the same variable in one equality constraint. The renaming is done by adding a version number to the variable. We call variables before and after renaming the \emph{original variable} and the \emph{renamed variable}. For example, we can rename $v_1$ to $v_1^{(0)}$, $v_1^{(1)}$, etc. In this case, $v_1$ is the original variable and both $v_1^{(0)}$ and $v_1^{(1)}$ are the renamed variables. For an original variable $v$, we use $R_v$ to denote the set of variables renamed from $v$. Later we will add arithmetic constraints to enforce that variables obtained from the same original variable are assigned the same value.

@@

$$ \Psi_r \defeq
\bigwedge_{v\in V_X, v'\in R_v} v=v' \wedge \#v = \#v'
$$
Here $V_X$ is the set of all character variables occurred in the PFAs of $x_1,x_2,\ldots,x_m$, i.e., $V_X=\bigcup_{i\in [1,m]}{V_{x_i}}$.

@@

\begin{lemma}\label{lemma:eqcorrect}
	$\semof {\syncfof A {A'}  \wedge \Psi_r}_{V_X \cup \#V_X} = \{\bigcup_{i\in[1,m]}(I^i\cup\pim^i)\mid w_1\cdot w_2 \cdots w_n = w_{n+1}\cdot w_{n+2} \cdots w_m \wedge \bigwedge_{i\in [1,m]}w_i=\decode A_{x_i}(I^i\cup\pim^i) \}$. 
\end{lemma}

>>>>

\newcommand{\nat}{\mathbb{N}}
\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\yfc}[1]{{\color{blue}YFC: #1}}
\newcommand{\petr}[1]{{\color{pink}Petr: #1}}
\newcommand{\sti}[1]{\mbox{\textsf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
%\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvarone}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvartwo}{\mbox{$V'_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{V}
\newcommand{\cvartwo}{V'}
\newcommand{\cvar}{V}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}[1]{\mathit{decode_{#1}}}
%\newcommand{\parikhfof}[1]{\Phi_{\parikhof{#1}}}
\newcommand{\parikhfof}[1]{\Phi_{\mathbb{P}}(#1)} %shouldn't it be something that takes automaton as parameter and returns formula?
\newcommand{\pim}{I_{\#}}
%\newcommand{\syncop}{\Cap}
\newcommand{\syncop}{\times}
\newcommand{\syncof}[2]{#1 \syncop #2}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncT}{T_\syncop}
\newcommand{\pvarsof}[1]{\#{#1}}
\newcommand{\pvar}{\pvarsof V}
\newcommand{\pvarone}{\pvarsof V}
\newcommand{\pvartwo}{\pvarsof {V'}}
\newcommand{\defeq}{::=}

@@

Through the paper, we will use quantifier-free linear integer arithmetic formulas, and call them \emph{linear formulas} for short.   
Given a linear formula $\phi$, a set of variables $V$ \lh{remove this?: that includes all variables occurring in $\phi$}, and an \emph{interpretation} of $V$, i.e., a function $I:V\rightarrow \nat$, 
we denote by $I\models_V \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner), and call $I$ a \emph{model} of $\phi$ over $V$. \lh{Our use of indexing by $V$ in the text later is not very nice. Add this for clarity?: If $V$ contain additional variables not present in $\phi$, the these might be assigned arbitrary values by $I$. If $V$ is missing some of the variables of $\phi$, then it must be possible to complete $I$ with assignment of the missing variables into a satisfying assignment of $\phi$ (as in existential quantification).}
We use $\modelsof \phi$ to denote the set all models of $\phi$ over $V$. 
We drop the reference to $V$ and write just ``model of $\phi$'',  $\modelsof \phi$, and $\models$ when no confusion may arise. For a set of interpretations $S$, we use $S_\Sigma$ to denote the set of interpretations obtained from $S$ by projecting the domain to $\Sigma$.

@@

The crucial feature of parametric flat automata is that their semantics can be faithfully represented by a linear formula and handled efficiently by an SMT solver. 
\lh{, allowing to circumvent more costly automata algorithms and decision procedures. Concretize and justify this better?}\yfc{The two sentences below read a bit strange, maybe need to rephrase it.}\lh{made an attempt}
%
This is facilitated by the structural constraints of PFA, which guarantee that 
every parametric word $x\in L(A)$ is uniquely determined by its Parikh image $\parikhof{x}$, as formalized by the following lemma.

\begin{lemma}\label{lemma:parikh}
For an PFA $A$ and $x,y\in L(A)$, if $\parikhof x = \parikhof y$ then $x = y$. 
\end{lemma}
%
Indeed, since every variable appears on at most one transition, then the $\parikhof x$ value of all variables appearing within the same cycle is the same, and it is equal to the number of repetitions of that cycle in the accepting run. 
%
The accepting run on $x$ (and so $x$ itself) can thus be reconstructed from $\parikhof x$. 
%
For example, in the automaton of Figure~\ref{fig:sfa_def}, 
from $\parikhwof x {v_0} = \parikhwof x {v_1} = \cdots = \parikhwof x {v_{m-1}}=1$ and $\parikhwof x {v^0_1}=\parikhwof x {v^1_1}=\parikhwof x {v^2_1}=2$ we derive that $x=v_0(v^0_1v^1_1v^2_1)^2v_1\cdots v_{m-1}$. 
We may say that $\parikhof x$ is an encoding of $x$, and define the decoding function $\decode A$ such that $\decode A(\parikhof x) = x$.
%

\begin{lemma}\label{lemma:decx}
For a PFA $A$, $L(A) = \{\decode A(\pim) \mid \pim\in\semof{\parikhfof A}\}$.
\lh{should we replace $\semof{\parikhfof A}$ by $\parikhof{A}$ here and almost everywhere?}
\end{lemma}
%
\begin{proof}
The function $\decode A$ can particularly be implemented as follows. Given $\pim:\pvars\rightarrow\nat$,
and assuming that the lengths of the loops of $A$ are $l_0,\ldots,l_{m-1}$, 
$\decode A(\pim)$ is constructed as the word $x_0 v_{0} \ldots v_{m-1} x_{m-1}$ where for each $i\in [0,m-1]$,
$x_i = (v_i^0 \cdots v_i^{l_{i-1}})^{\pvarsof {v_i^0}}$ if $l_i >0$ and $x_i = \epsilon$ if $l_i = 0$. 
\end{proof}

Since every word $w \in \semof{A}$ is obtained from a parametric word $x\in L(A)$ and an interpretation of character variables $I:\cvar\rightarrow\Sigma_\epsilon$ such that $w = I(x)$, and since $x = \decode A(\parikhof x)$, 
we may say that the assignment $I \cup \parikhof x$ uniting the interpretation $I$ and the Parikh image of the parametric word $x$ encodes $w$. Lemma~\ref{lemma:parikh} also implies that two different words are encoded by different united assignments (the Parikh image or the interpretation must be different), and hence we may extend the decoding function $\decode A$ to united assignments so that $\decode A(I \cup \parikhof x) = w$.
Encodings of words in $\semof{A}$ are encodings of parametric words of $L(A)$ paired with arbitrary interpretations of character variables $\cvar$, which are precisely all models of the formula $\parikhfof A$ interpreted over $\cvar\cup\pvarsof \cvar$, as formulated by the following corollary of Lemma~\ref{lemma:decx}:
%
\begin{corollary}\label{corollary:pfa}
For an PFA $A$ over character variables $V$,
$\semof{A} = \{\decode A(I\cup \pim) \mid I\cup\pim\in\semof{\parikhfof A}_{\cvar\cup\pvar}\}$.
%$\semof{A} = \{I(\decode A(\pim)) \mid \pim\in\semof{\parikhfof A},I:\cvars\rightarrow\Sigma_\epsilon\}$.
\end{corollary}
%
A minor technical difficulty is that we wish to work with pure arithmetic formulas, 
but we may also wish to specify when a character variable is assigned $\epsilon$, which is not a number. 
Hence, to allow purely arithmetic formulas to talk about the value $\epsilon$, we encode $\epsilon$ as $-1$.
In the further development, $\epsilon$ will mean $-1$ when working with arithmetic formulas constraining interpretations of character variables.

\paragraph{Synchronization of PAs}
\ \lh{changed the formulation of the next two paragraphs a bit, is it ok?}
We will now discuss a construction of the \emph{synchronization formula} for two PAs $A$ and $A'$. 
%over character variables $V$ and $V'$. 
It is a linear formula $\syncfof A {A'}$ 
%over $\cvarone\cup \cvartwo \cup \pvarone\cup\pvartwo$ 
that specifies how each word in the semantic intersection $\semof A \cap \semof {A'}$ is encoded by $A$ and by $A'$.
%
More precisely, the models of $\syncfof A {A'}$ represent pairs of word
encodings $(I\cup \parikhof{x})$ and $(I'\cup \parikhof{x'})$ such that $x\in L(A)$, $x'\in L(A')$, and 
$I(x) = I'(x')$.
% \in \semof{A}\cap\semof{A'}$.
The synchronization formula will be used in, e.g., the decision procedure of equality constraints $t_1=t_2$ to find a common interpretation of $t_1$ and $t_2$.  
%More precisely, the models of the synchronization formula $\syncfof A {A'}$ correspond to the set of all pairs of word
%encodings $(I\cup \parikhof{x})$ and $(I'\cup \parikhof{x'})$ such that $x\in L(A)$, $x'\in L(A')$, and 
%$I(x) = I'(x') \in \semof{A}\cap\semof{A'}$.%, which constitute the intersection.
% encodings $(I\cup \pim)\in \parikhof {A}$ and $(I'\cup \pim')\in\parikhof {A'}$ such that 
%$I(\decode A(I\cup \pim) = \decode {A'}(I'\cup \pim')$, which constitute the intersection.

Particularly, the synchronization formula is build for two PAs  
$A = (Q,T,\cvarone,q_i,q_f)$ and ${A'} = (Q',T',\cvartwo,q_i',q_f')$ such that $\cvarone \cap \cvartwo = \emptyset$. 
%
It is extracted from the \emph{asynchronous product} of $A$ and $A'$. 
%
The asynchronous product is an automaton that uses $Q\times Q'$ as the set of states and 
$\cvarone_\epsilon\times\cvartwo_\epsilon$
as the alphabet. 
%
Every accepting run of $\syncof A {A'}$ corresponds to a pair of accepting runs, 
a run of $A$ over a parametric word $x$ and a run of ${A'}$ over a parametric word $x'$. 
The word accepted by the run of $\syncof A {A'}$ induces constraints on the interpretations of $I$ of $\cvarone$ and $I'$ of $\cvartwo$ under which the two parametric words have the same interpretation, i.e. $I(x) = I'(x')$.

@@

The correctness of this construction is stated in Lemma~\ref{lemma:synccorrect} below. The lemma can also be understood as a precise specification of the semantics of the synchronization formula, on which we rely for correctness of the construction of under-approximations of string constraints in Sections~\ref{section:mem}, \ref{section:eq}, and \ref{section:s2i}. 
%
\hide{
\begin{lemma}\label{lemma:synccorrect}
%$\semof {\exists V_\times . \syncfof A {A'}}\restrict{\cvarone \cup \cvartwo \cup \pvarsof{\cvarone} \cup \pvarsof\cvartwo} = \{(I\cup\pim)\cup(I'\cup\pim') \mid \decode A(I\cup\pim) = \decode {A'}(I'\cup\pim')\}$. 
$\semof {\syncfof A {A'}}_{\cvarone\cup \cvartwo \cup \pvarone\cup\pvartwo} = \{(I\cup\pim)\cup(I'\cup\pim') \mid \decode A(I\cup\pim) = \decode {A'}(I'\cup\pim')\}$. 
\end{lemma}
%
Another version:}
\begin{lemma}\label{lemma:synccorrect}
$
\semof {\syncfof A {A'}}_{\cvarone\cup \cvartwo \cup \pvarone\cup\pvartwo} = 
\{
(I\cup\parikhof{x})\cup(I'\cup\parikhof{x'}) \mid 
x\in L(A), x'\in L(A'), I(x) = I(x')
\}
$
\end{lemma}
Informally, the lemma states that the models of $\syncfof A {A'}$ encode precisely the pairs of equivalent encodings of words from $\semof{A}$ and $\semof{A'}$, that constitute the intersection $\semof{A} \cap \semof{A'}$. A minor technical difficulty is that the models of $\syncfof A {A'}$ include also an assignment to the auxiliary variables of $(\cvarone_\epsilon\times\cvartwo_\epsilon) \cup \pvarsof{(\cvarone_\epsilon\times\cvartwo_\epsilon)}$. Hence in the lemma, we have to restrict the models to the original variables $\cvarone \cup \cvartwo \cup \pvarsof{\cvarone} \cup \pvarsof\cvartwo$. %This is  achieve by quantifying out the variables of $V_\times$.

Notice that if $A$ is flat (or, symmetrically, if $A'$ is flat), then the semantic intersection $\semof A \cap \semof {A'}$ can be still decoded from the synchronisation formula $\semof {\syncfof A {A'}}$. 
Namely, due to Corollary~\ref{corollary:pfa}, we have that $\semof A \cap \semof {A'} = \{\decode A(I\cup \pim) \mid (I\cup \pim) \in \semof{\syncfof A {A'}}_{\cvarone\cup \pvarone}\}$.

%We actually have even something stronger, namely, that, $\syncfof A {A'}$ has all possible allignments. How do I say this ... or do I have to say this?
%\begin{lemma}
%For $\pim:\pvarone\rightarrow \nat$, $\pim':\pvartwo \rightarrow\nat$, $I:\cvarone\rightarrow\Sigma$, $I':\cvartwo\rightarrow\Sigma$, 
%we have $(I\cup\pim)\cup(I'\cup\pim')\in\semof {\syncfof A {A'}}$ if and only if $\decode A(I\cup\pim) = \decode {A'}(I'\cup\pim')$.
%\end{lemma}

\lh{this following sentence appears to be in a wrong place here. Would be nicer in some introductory part to the underapproximation which talks about this, but I don't know where it should be.}
In the rest of the section, we assume each string variable $x$ is restricted to a PFA $A_x$ with the set of character variables~$V_x$.

\yfc{we need to define \#{V} somewhere}
\lh{it is in prelims next to Parikh im. definition}

@@

\begin{itemize}
	\item $V' = \{ v_{(q,a,q')} \mid (q,a,q')\in T\}$
	\item $T' = \{ (q, v_{(q,a,q')}, q')  \mid (q,a,q')\in T\}$
\end{itemize}
We use the linear formula $\phi_c \defeq \bigwedge\limits_{(q,a,q')\in T} (v_{(q,a,q')} = a)$ to remember the relation between the fresh character variables in $V'$ and the symbol it represents. The relation between $A$ and $A'$ is expressed in Lemma~\ref{lemma:fa_pfa}.

\begin{lemma}\label{lemma:fa_pfa}
	$L(A) =  \{  \decode {A'}(I\cup\pim) \mid (I\cup\pim)\in  {\semof {\parikhfof {A'} \wedge \phi_c}} _{V' \cup \#V'} \}$. 
\end{lemma}

Intuitively, we create a new character variable for each transition in $T$, later we will add the relation $v_{(q,a,q')}  = a$ to $\Phi_O$.

The correctness of this construction is expressed in Lemma~\ref{lemma:memcorrect} below. 
%
%\newcommand\restrict[1]{\raisebox{-.5ex}{$\big|$}_{#1}}
\begin{lemma}\label{lemma:memcorrect}
	$({\semof {\syncfof {A_x }{A'}} \cap \bigcap_{\phi\in\Phi_C} \semof {\phi} } )_{V_x \cup \#V_x} = \{(I\cup\pim) \mid \decode {A_x}(I\cup\pim) \in L(A)\}$. 
\end{lemma}

\lh{Simplified version:
Assume that the constraint we take from $\Phi_I$ is a regular constraint $\phi_r \defeq x \in L(A)$. 
%Assume that $A=(Q,T,\Sigma,q_i,q_f)$. 
We represent the FA $A$ by a pair of a PA $A'$ and linear formula $\phi_A$. 
The PA $A'$ operates over fresh character variables $v_a, a\in\Sigma_\epsilon$, and is obtained from $A$ by replacing every occurrence of each character $a\in\Sigma_{\epsilon}$ on a transition by the variable $v_a$. 
The formula $\phi_A$ then binds the fresh variables to the characters they represent, 
namely, $\phi_A = \bigwedge_{a\in\Sigma_\epsilon} v_a = a$.
The resultinng overapproximation of $\phi_r$ is then the formula $\syncfof{A_x} {A'} \land \phi_A$.

The following lemma states the correctness of this construction: Intuitively, the resulting constraint represents presicely the restriction of $L(A)$ to the domain $\semof{A_x}$, the $A_x$-encodings of the strings in $L(A)\cap\semof{A_x}$. 
%
\begin{lemma}\label{lemma:memcorrect}
	${\semof {\syncfof {A_x }{A'} \land \phi_A} } _{V_x \cup \pvarsof{V_x}} = \{(I\cup\pim) \mid \decode {A_x}(I\cup\pim) \in L(A)\}$. 
\end{lemma}
}

@@

Assume that the constraint we take from $\Phi_I$ is an equality constraint $\phi_e\defeq x_1\cdot x_2 \cdots x_n = x_{n+1}\cdot x_{n+2} \cdots x_m$. We first build two PFAs that encode the left-hand side term and the right-hand side term of the equality constraint. Intuitively, this can be achieved by connecting the PFAs of the string variables with epsilon transitions. However, there are two minor technical issues. First, our definition of PFA does not allow epsilon transition. Second, some string variable might occur multiple times in a equality constraint, but the definition of PFA allows each character variable occurs in only one transition. \lh{the second is maybe not a minor technical issue}

@@

For the second problem, we rename the variables on the transitions of PFAs for different occurrences of the same variable in one equality constraint. The renaming is done by adding a version number to the variable. We call variables before and after renaming the \emph{original variable} and the \emph{renamed variable}. For example, we can rename $v_1$ to $v_1^{(0)}$, $v_1^{(1)}$, etc. In this case, $v_1$ is the original variable and both $v_1^{(0)}$ and $v_1^{(1)}$ are the renamed variables. For an original variable $v$, we use $R_v$ to denote the set of variables renamed from $v$. Later we will add arithmetic constraints to enforce that variables obtained from the same original variable are assigned the same value. 
\lh{I did not find those, are they somewhere or forgoten?}

@@

$$ \Psi_r \defeq
\bigwedge_{v\in V_X, v'\in R_v} v=v' \wedge \#v = \#v'
$$
Here $V_X$ is the set of all character variables occurred in the PFAs of $x_1,x_2,\ldots,x_m$, i.e., $V_X=\bigcup_{i\in [1,m]}{V_{x_i}}$. \lh{what is $R_v$?}

@@

\begin{lemma}\label{lemma:eqcorrect}
	$\semof {\syncfof A {A'}  \wedge \Psi_r}_{V_X \cup \#V_X} = \{\bigcup_{i\in[1,m]}(I^i\cup\pim^i)\mid w_1\cdot w_2 \cdots w_n = w_{n+1}\cdot w_{n+2} \cdots w_m \wedge \bigwedge_{i\in [1,m]}w_i=\decode A_{x_i}(I^i\cup\pim^i) \}$.\lh{aaarrrghhh ... } 
\end{lemma}
